---
title: "Peer Assessment 1"
author: "Aníbal Brenes Jiménez"
date: "16 July 2015"
output: html_document
---

##Loading and preprocessing the data

```{r load data}
dat<-read.csv("C:/Users/anbren2/Documents/Data_Science/05 Reproducible research/activity.csv")
head(dat)
```
##Mean total number of steps taken per day

```{r mean}
daytotal <- 1:(nrow(dat)/288)
for (i in 1:length(daytotal)){
  x <- (i-1)*288+1
  y <- 288*i
  daytotal[i] <- sum(dat$steps[x:y],na.rm = T)
}

daytotal

hist(daytotal,breaks = seq(from = 0, to = max(daytotal,na.rm = T), by = max(daytotal,na.rm = T)/10 ), xlab = "Number of steps", main = "Histogram with raw data")

meandt <- mean(daytotal,na.rm = T)
meandt  <- round(meandt, digits = 2)
mediandt <- median(daytotal,na.rm = T)

```
The mean of the total number of steps taken per day is `r meandt` and the median is `r mediandt`

##Average daily activity pattern

```{r}
interv <- dat$interval[1:288]
intmean <- NULL
intsum <- NULL
for (i in 1:288){
  intmean[i]<-mean(dat$steps[dat$interval==interv[i]],na.rm=T)
  }
plot(interv,intmean,type = "l")
maxsteps<-interv[intmean == max(intmean)]
```

The interval which, on average, contains the maximum number of steps is `r maxsteps`


#Imputing missing values


```{r}
qna <- sum(is.na(dat$steps))
```


The total number of missing values in the dataset is `r qna`

```{r}
#Missing values will be filled with the mean for the corresponding 5-min interval
meancomp <- rep(intmean, times = 61)
datcomp <- dat
datcomp$steps[is.na(datcomp$steps)]<-meancomp[is.na(datcomp$steps)]

daytotal <- 1:(nrow(dat)/288)
for (i in 1:length(daytotal)){
  x <- (i-1)*288+1
  y <- 288*i
  daytotal[i] <- sum(datcomp$steps[x:y])
}
daytotal
hist(daytotal,breaks = seq(from = 0, to = max(daytotal), by = max(daytotal)/10 ), xlab = "Amount of steps", main = "Histogram with filled data")

meandt <- mean(daytotal)
meandt  <- round(meandt, digits = 0)
mediandt <- median(daytotal)
mediandt  <- round(mediandt, digits = 0)
```
The mean of the total number of steps taken per day, after filling the missing values is `r meandt` and the median is `r mediandt`.

The mean value has been increased considerably due to in the first case, days without data are considered to have 0 steps (that is possible to see in the first histogram), and now these days have 10766.19 steps. Besides, the median was also increased for the same reason.  

##Differences in activity patterns between weekdays and weekends

```{r weekdays}
library(timeDate)
wdays <- isWeekday(as.POSIXct(dat$date))
datcomp <- data.frame(datcomp, day = wdays)
datcomp$day[datcomp$day==T] <- "Weekday"
datcomp$day[datcomp$day==F] <- "Weekend"
head(datcomp)

intmeanwd <- NULL
intmeanwe <- NULL

for (i in 1:288){
intmeanwd[i]<-mean(datcomp$steps[datcomp$interval==interv[i]& datcomp$day == "Weekday" ],na.rm=T)
intmeanwe[i]<-mean(datcomp$steps[datcomp$interval==interv[i]& datcomp$day == "Weekend" ],na.rm=T)
}

par(mfrow=c(3,1))
plot(interv,intmeanwd,type="l", main = "Weekdays", xlab = "5 min interval", ylab = "Avg steps taken")
plot(interv,intmeanwe,type="l",col = "red", main = "Weekends",xlab = "5 min interval", ylab = "Avg steps taken")
plot(interv,intmeanwd,type="l",xlab = "5 min interval", ylab = "Avg steps taken")
lines(interv,intmeanwe, col="red")

```

